---
title: "Project 1 Vignette"
author: "DRV"
date: "9/18/2020"
output: 
rmarkdown::github_document:
  toc: TRUE 
  toc_depth: 6 
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Packages required to create this vignette: 
httr, jsonlite, dplyr, ggplot2, knitr, rmarkdown

### Contacting API
```{r ContactAPIs, include=TRUE}
library(httr)
library(jsonlite)
library(dplyr)

### Franchise Data 
url1 <- "https://records.nhl.com/site/api/franchise"
call1 <- paste(url1)
get_franchise <- GET(call1)
get_franchise_text <- content(get_franchise, "text")
get_franchise_json <- fromJSON(get_franchise_text, flatten=TRUE)
get_franchise_df <- as.data.frame(get_franchise_json)

### Franchise team Totals 
url2 <- "https://records.nhl.com/site/api/franchise-team-totals"
call2 <- paste(url2)
get_ftt <- GET(call2)
get_ftt_text <- content(get_ftt, "text")
get_ftt_json <- fromJSON(get_ftt_text, flatten=TRUE)
get_ftt_df <- as.data.frame(get_ftt_json)

### All Time Record vs. Franchise 
url3 <- "https://records.nhl.com/site/api/all-time-record-vs-franchise?cayenneExp=teamFranchiseId="
function1 <- function(team){paste(url3,team,sep="")}

team <- "Bruins" #Insert Team Common Name or ID here 

call3 <- function1(ifelse(is.integer(team), team, switch(team, "Canadians" = 8, "Wanderers" = 41, "Eagles" = 45, "Tigers" = 37, "Maple Leafs" = 10, "Bruins" = 6, "Maroons" = 43, "Americans" = 51, "Quakers" = 39, "Rangers" = 3, "Blackhawks" = 16, "Red Wings" = 17, "Barons" = 49, "Kings" = 26, "Stars" = 25, "Flyers" = 4, "Penguins" = 5, "Blues" = 19, "Sabres" = 7, "Canucks" = 23, "Flames" = 20, "Islanders" = 2, "Devils" = 1, "Capitals" = 15, "Oilers" = 22, "Hurricanes" = 12, "Avalanche" = 21, "Coyotes" = 23, "Sharks" = 28, "Senators" = 9, "Lightening" =14, "Ducks" = 24, "Panthers" = 13, "Predators" = 18, "Jets" = 52, "Blue Jackets" = 29, "Wild" = 30, "Golden Knights" = 54)))

get_atrvf <- GET(call3)
get_atrvf_text <- content(get_atrvf, "text")
get_atrvf_json <- fromJSON(get_atrvf_text, flatten=TRUE)
get_atrvf_df <- as.data.frame(get_atrvf_json)

### Franshice Goalie Records 
url4 <- "https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId="
function2 <- function(team2){paste(url4,team2,sep="")}

team2 <- "Canadians" #Insert Team Common name or ID here 

call4 <- function2(ifelse(is.integer(team2), team2, switch(team2, "Canadians" = 8, "Wanderers" = 41, "Eagles" = 45, "Tigers" = 37, "Maple Leafs" = 10, "Bruins" = 6, "Maroons" = 43, "Americans" = 51, "Quakers" = 39, "Rangers" = 3, "Blackhawks" = 16, "Red Wings" = 17, "Barons" = 49, "Kings" = 26, "Stars" = 25, "Flyers" = 4, "Penguins" = 5, "Blues" = 19, "Sabres" = 7, "Canucks" = 23, "Flames" = 20, "Islanders" = 2, "Devils" = 1, "Capitals" = 15, "Oilers" = 22, "Hurricanes" = 12, "Avalanche" = 21, "Coyotes" = 23, "Sharks" = 28, "Senators" = 9, "Lightening" =14, "Ducks" = 24, "Panthers" = 13, "Predators" = 18, "Jets" = 52, "Blue Jackets" = 29, "Wild" = 30, "Golden Knights" = 54)))

get_fgr <- GET(call4)
get_fgr_text <- content(get_fgr, "text")
get_fgr_json <- fromJSON(get_fgr_text, flatten=TRUE)
get_fgr_df <- as.data.frame(get_fgr_json)

### Franchise Skater Records 
url5 <- "https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId="

team3 <- "Bruins" #Insert Team Common name or ID here

function3 <- function(team3){paste(url5,team3, sep="")}

call5 <- function3(ifelse(is.integer(team3), team3, switch(team3, "Canadians" = 8, "Wanderers" = 41, "Eagles" = 45, "Tigers" = 37, "Maple Leafs" = 10, "Bruins" = 6, "Maroons" = 43, "Americans" = 51, "Quakers" = 39, "Rangers" = 3, "Blackhawks" = 16, "Red Wings" = 17, "Barons" = 49, "Kings" = 26, "Stars" = 25, "Flyers" = 4, "Penguins" = 5, "Blues" = 19, "Sabres" = 7, "Canucks" = 23, "Flames" = 20, "Islanders" = 2, "Devils" = 1, "Capitals" = 15, "Oilers" = 22, "Hurricanes" = 12, "Avalanche" = 21, "Coyotes" = 23, "Sharks" = 28, "Senators" = 9, "Lightening" =14, "Ducks" = 24, "Panthers" = 13, "Predators" = 18, "Jets" = 52, "Blue Jackets" = 29, "Wild" = 30, "Golden Knights" = 54)))

get_fsr <- GET(call5)
get_fsr_text <- content(get_fsr, "text")
get_fsr_json <- fromJSON(get_fsr_text, flatten=TRUE)
get_fsr_df <- as.data.frame(get_fsr_json)

### NHL Stats Data 
url6 <- "https://statsapi.web.nhl.com/api/v1/teams/"
function4 <- function(team4, modifier){paste(url6, team4, modifier, sep="")}

team4 <- "Bruins" #Insert Team Common name or ID here

call6 <- function4(ifelse(is.integer(team3), team4, switch(team4, "Canadians" = 8, "Wanderers" = 41, "Eagles" = 45, "Tigers" = 37, "Maple Leafs" = 10, "Bruins" = 6, "Maroons" = 43, "Americans" = 51, "Quakers" = 39, "Rangers" = 3, "Blackhawks" = 16, "Red Wings" = 17, "Barons" = 49, "Kings" = 26, "Stars" = 25, "Flyers" = 4, "Penguins" = 5, "Blues" = 19, "Sabres" = 7, "Canucks" = 23, "Flames" = 20, "Islanders" = 2, "Devils" = 1, "Capitals" = 15, "Oilers" = 22, "Hurricanes" = 12, "Avalanche" = 21, "Coyotes" = 23, "Sharks" = 28, "Senators" = 9, "Lightening" =14, "Ducks" = 24, "Panthers" = 13, "Predators" = 18, "Jets" = 52, "Blue Jackets" = 29, "Wild" = 30, "Golden Knights" = 54)), "?expand=team.roster")

get_stats <- GET(call6)
get_stats_text <- content(get_stats, "text")
get_stats_json <- fromJSON(get_stats_text, flatten=TRUE)
get_stats_df <- as.data.frame(get_stats_json)
  
### Wrapper Function need to add name coercer 

function5 <- function(url, team5, modifier){
  call7 <- paste(url, ifelse(is.integer(team5), team5, switch(team5, "Canadians" = 8, "Wanderers" = 41, "Eagles" = 45, "Tigers" = 37, "Maple Leafs" = 10, "Bruins" = 6, "Maroons" = 43, "Americans" = 51, "Quakers" = 39, "Rangers" = 3, "Blackhawks" = 16, "Red Wings" = 17, "Barons" = 49, "Kings" = 26, "Stars" = 25, "Flyers" = 4, "Penguins" = 5, "Blues" = 19, "Sabres" = 7, "Canucks" = 23, "Flames" = 20, "Islanders" = 2, "Devils" = 1, "Capitals" = 15, "Oilers" = 22, "Hurricanes" = 12, "Avalanche" = 21, "Coyotes" = 23, "Sharks" = 28, "Senators" = 9, "Lightening" =14, "Ducks" = 24, "Panthers" = 13, "Predators" = 18, "Jets" = 52, "Blue Jackets" = 29, "Wild" = 30, "Golden Knights" = 54)), modifier, sep="") 
  get_data <- GET(call7) 
  get_data_text <- content(get_data, "text") 
  get_data_json <- fromJSON(get_data_text, flatten=TRUE) 
  get_data_df <- as.data.frame(get_data_json)
  }

### Testing Wrapper Function
get_data_df <- function5("https://statsapi.web.nhl.com/api/v1/teams/","Hurricanes","?expand=team.roster")

```
## Exploratory Data Analysis 
### Joining Franchise Team Totals and Franchise Goalie Records on "data.franchiseId" 
```{r EDA, include=TRUE}
joined_data <- inner_join(get_ftt_df, get_fgr_df, by = c("data.franchiseId" =  "data.franchiseId"))
```

### Creating two new variables 
```{r NewVars, include=TRUE}

get_fgr_df <- mutate(get_fgr_df, DepthChart= ifelse(get_fgr_df$data.gamesPlayed > 100, "Starter", "Backup")) 

get_fgr_df <- mutate(get_fgr_df, CareerWP500 = ifelse((get_fgr_df$data.gamesPlayed-get_fgr_df$data.losses)/data.gamesPlayed > .5, "Over500", "Under500"))

```
### Contingency tables 
The first table shows that there is only one active starter and one active backup grouped goalie, and that there are many more backup rated goalies that are inactive when compared to starter rated goalies. 
```{r Cont1, include=TRUE}
library(knitr)
firstctable <- as.data.frame(table(get_fgr_df$data.activePlayer, get_fgr_df$DepthChart))
kable(firstctable, caption = "Contingency Table for Player Status(Active/Not Active) vs. Experience More/Less than 100 games(Backup/Starter")
```
The second table shows that the two active goalies for the canadians have a career Win Percentage of over 50% and that most of the goalies that they have had which are now inactive, also mostly all have career Win Percentage > 50%. 
```{r Cont 2, include=TRUE}
secondctable <- as.data.frame(table(get_fgr_df$data.activePlayer, get_fgr_df$CareerWP500))
kable(secondctable, caption = "Contingency table for Games Played vs. Career WP > .500")
```
### Numerical Summaries for Numerical Variables grouped by Categorical Variables
The first numerical summary table shows the average # of wins was much greater for the Starter categorical variable compared to the backup. 
```{r NumericalSummaries, include=TRUE}
ntable <- get_fgr_df %>% group_by(DepthChart) %>% summarise(avg = mean(data.wins), med = median(data.wins), var = var(data.wins))

kable(ntable)
```
The second numerical summary table shows that the average # of shutouts was much higher for goalies with a career Win percentages above 50%. 
```{r NumericalSummary2,include=TRUE}
ntable2 <- get_fgr_df %>% group_by(CareerWP500) %>% summarise(avg = mean(data.mostShutoutsOneSeason), med = median(data.mostShutoutsOneSeason), var = var(data.mostShutoutsOneSeason))

kable(ntable2)
```

### Bar Plot
This bar plot shows the max saves in one game of all of the goalies that have played for the Canadians. Above 50: Niemi, Plante, Price, Thomas. These goalies can be inferred to have played a great singular game in terms of the goalie position.
```{r Bar, include=TRUE}
library(ggplot2)
ggplot(get_fgr_df, aes(x = data.lastName, y = data.mostSavesOneGame, color = data.lastName, fill=data.lastName)) + theme(axis.text.x=element_text(angle=90,hjust=1),legend.position = "none") + geom_col()+ggtitle("Montreal Canadians Goalies - Most Saves In One Game") + xlab("Goalie Last Name") + ylab("Saves in One Game") 

```
### Histogram
This histogram shows that the majority of the canadians goalies had a max of between 4 and 8 goals scored against them. 
```{r Histogram, include=TRUE}
ggplot(get_fgr_df, aes(x = data.mostGoalsAgainstOneGame)) + geom_histogram(bins = 10) + ggtitle("Histogram for Most Goals Against in One Game") 
```

### Box Plot 
This boxplot shows that the majority of canadians goalies who have a career win percentage above 50% played a lot more games than the goalies that had a career win percentage of below 50%. 
```{r BoxPlot, include=TRUE}
library(ggplot2)
ggplot(get_fgr_df, aes(x = CareerWP500, y = data.gamesPlayed)) + geom_boxplot() + 
  geom_jitter(aes(color = CareerWP500)) + ggtitle("Canadians Goalies Total Games Stratified by Win Percentage") + xlab("Over or below .500 Win %") + ylab("Number of games played in Career")
```

# Scatter Plot 
This scatter plot and the associated linear model shows that for the goalies with a greater than 50% career win percentage, the more games you play, the more your win % goes down. 
```{r Scatter, include=TRUE}
ggplot(get_fgr_df, aes(x = data.gamesPlayed, y =((data.gamesPlayed-data.losses)/data.gamesPlayed) , group = CareerWP500)) + geom_point(aes(color = CareerWP500)) +     
   geom_smooth(method = 'lm', color = 'green') + ggtitle("Canadians Goalies Games played vs. Win Percentage")+ xlab("Games Played in Career") + ylab("Win %")
```

# Final Plot 
This scatter plot shows that here is somewhat of a correlation between Penalty minutes and points, but this interaction can probably be attributed mostly to another variable such as playing time. 
```{r Final, include=TRUE}
ggplot(get_fsr_df, aes(x = data.penaltyMinutes, y = data.points, group = data.positionCode)) + geom_point(aes(color = data.positionCode)) +     
   geom_smooth(method = 'lm', color = "red") + ggtitle("Penalty Minutes vs. Points")+ xlab("Penalty Minutes") + ylab("Points")

```
